---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# CellPhe

<!-- badges: start -->
<!-- badges: end -->

CellPhe provides functions to accompany the paper (TODO Add reference when available) to phenotype cells from time-lapse videos.

## Installation

You can install the latest version of CellPhe from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("uoy-research/CellPhe")
```

## Example

Included with the package is an example dataset to demonstrate CellPhe's capabilities, this data is available in `data` and comprises 3 parts:

  - The time-lapse stills as TIFF images (`data/05062019_B3_3_imagedata`)
  - Existing pre-extracted features (`data/05062019_B3_3_Phase-FullFeatureTable.csv`)
  - Region-of-interest boundaries already demarked (`data/05062019_B3_3_Phase.zip`) 

```{r example}
library(CellPhe)
trial_name = "05062019_B3_3"
basedir <- "data"
```

The first step is to read in the raw images and normalise the pixel values to [0, 255].

```{r}
imagedata =  sprintf("%s/%s_imagedata", basedir, trial_name)
frames = readTiffs(imagedata)
normalised_frames = lapply(frames, normaliseImage, lower = 0, upper = 255)
```

Optionally, if there are already features available for each cell tracked they can be loaded in and copied.
This step isn't obligatory.
(TODO What format should this CSV be in? 1 row per frame, with at least FrameId and CellId columns?)

```{r}
input_feature_table =  sprintf("%s/%s_Phase-FullFeatureTable.csv", basedir, trial_name)
min_frames = 50
feature_table = copyFeatures(input_feature_table, min_frames)
numcells_over_thresh = feature_table[[1]]
original_IDs = feature_table[[2]]
missing_frames = feature_table[[3]]
features = feature_table[[4]]
```

Given a directory containing ROI files (TODO what format should these be in?), the `extractFeatures` function can be used to extract time-series features for each cell.

```{r}
roi_files =  sprintf("%s/%s_Phase", basedir, trial_name)
new_features = extractFeatures(roi_files, original_IDs, missing_frames, normalised_frames, min_frames)
```

The original features and new features can be combined together in the `varsFromTimeSeries` functions to extract a final list of variables.
TODO What exactly is this doing differently to extractFeatures?

```{r}
tsvariables = varsFromTimeSeries(features, new_features, trial_name, original_IDs)
```

TODO classification/clustering