---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# CellPhe

<!-- badges: start -->
<!-- badges: end -->

CellPhe provides functions to accompany the paper (TODO Add reference when available) to phenotype cells from time-lapse videos.

## Installation

You can install the latest version of CellPhe from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("uoy-research/CellPhe")
```

## Example

Included with the package is an example dataset to demonstrate CellPhe's capabilities, this data is available in `data` and comprises 3 parts:

  - The time-lapse stills as TIFF images (`data/05062019_B3_3_imagedata`)
  - Existing pre-extracted features (`data/05062019_B3_3_Phase-FullFeatureTable.csv`)
  - Region-of-interest (ROI) boundaries already demarked in ImageJ format (`data/05062019_B3_3_Phase.zip`) 

```{r example}
library(CellPhe)
trial_name <- "05062019_B3_3"
basedir <- "data"
```

The first step is to read in the raw images.

```{r}
imagedata <- sprintf("%s/%s_imagedata", basedir, trial_name)
frames <- readTiffs(imagedata)
```

Optionally, if there are already features available for each cell tracked 
they can be loaded in and copied. For example, the features requiring phase 
information can be copied in from the feature table output by PhaseFocus Livecyte 
software using the function `copyFeatures()`. Only the values for the features 
volume and sphericity, which require phase information, are copied as all other 
features can be calculated by the function `extractFeatures()`. Only cells that 
are tracked for a minimum of min_frames are copied into the new feature table. 
The output is a data frame with each row corresponding to a cell tracked in a
given frame. Only cells that are present in the specified minimum number of frames 
(`min_frames`). are included. 
The first 2 columns provide numerical identifiers for the frame and cell, while any
remaining columns are features extracted from the PhaseFocus feature table (currently just
volume and sphericity).

```{r}
# COPY PHASE FEATURES FROM PHASEFOCUS FEATURE TABLE FOR EACH CELL TRACKED FOR MORE THAN MIN_FRAMES AND FIND MISSING FRAMES:
min_frames <- 50
input_feature_table <- sprintf("%s/%s_Phase-FullFeatureTable.csv", basedir, trial_name)
feature_table <- copyFeatures(input_feature_table, min_frames)
```

Alternatively, pre-calculated features from any other tracking software can be provided
as long as it is organised in the same manner, with a dataframe with `FrameID` and `CellID`
as the first 2 columns and any features present in the remaining columns.

If no pre-existing features are available, a data frame with just `FrameID` and `CellID` columns
will need to be made available.

In addition to any pre-calculated features, the `extractFeatures()` function reads in information on cell 
boundaries (`roi_folder`) and together with the images for every frame (`frames`). 
The images and boundary information are used to calculate a total of 72 descriptive features for each cell 
on every frame based on size, shape, texture and movement including the local cell density 
The output is a dataframe comprising the `FrameID` and `CellID` columns, the 72 features as columns, and any additional features that may be present (such as from `copyFeatures()`) in further columns.

```{r}
roi_files <- sprintf("%s/%s_Phase", basedir, trial_name)
new_features <- extractFeatures(feature_table, frames, roi_files, min_frames, framerate=0.0028)
```

Variables are calculated from the time series for any pre-existing features as well as the output of `extractFeatures()`, providing both summary statistics and indicators of time-series behaviour at different levels of detail obtained via wavelet analysis. 
15 summary scores are calculated for each feature, in addition to the cell trajectory, thereby resulting in a default output of 1081 features (15x72 + 1).
These are output in the form of a dataframe with the first column being the `CellID` used previously.

```{r}
tsvariables = varsFromTimeSeries(new_features)
```

TODO classification/clustering
